---
title: "Assignment 6: GLMs (Linear Regressios, ANOVA, & t-tests)"
author: "Xuening Tang"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
#This code chunk will tidy your knit PDF files, wrapping long code lines
#For it to work, the "formatR" package needs to be installed

#install.packages('formatR')
#knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on generalized linear models. 

## Directions
1. Rename this file `<FirstLast>_A06_GLMs.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
5. When you have completed the assignment, **Knit** the text and code into a single PDF file.


## Set up your session 
1. Set up your session. Check your working directory. Load the tidyverse, agricolae and other needed packages. Import the *raw* NTL-LTER raw data file for chemistry/physics (`NTL-LTER_Lake_ChemistryPhysics_Raw.csv`). Set date columns to date objects.

2. Build a ggplot theme and set it as your default theme.

```{r setup}
#1
getwd()
library(tidyverse)
library(agricolae)
library(lubridate)
library(ggpmisc)
library(corrplot)
litter.chem.phy <- 
  read.csv("../Data/Raw/NTL-LTER_Lake_ChemistryPhysics_Raw.csv", stringsAsFactors = TRUE)
litter.chem.phy$sampledate <- 
  as.Date(litter.chem.phy$sampledate , format = "%m/%d/%y")

#2
my.theme <- theme_classic(base_size = 12) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(my.theme)
```

## Simple regression
Our first research question is: Does mean lake temperature recorded during July change with depth across all lakes?

3. State the null and alternative hypotheses for this question:
> Answer:
H0: Mean lake temperature recorded during July does not change with depth across all lakes.
Ha: Mean lake temperature recorded during July change with depth across all lakes.


4.  Wrangle your NTL-LTER dataset with a pipe function so that the records meet the following criteria: 
 * Only dates in July. 
 * Only the columns: `lakename`, `year4`, `daynum`, `depth`, `temperature_C`
 * Only complete cases (i.e., remove NAs)

5. Visualize the relationship among the two continuous variables with a scatter plot of temperature by depth. Add a smoothed line showing the linear model, and limit temperature values from 0 to 35 °C. Make this plot look pretty and easy to read.

```{r scatterplot}
#4
NTL.litter <- 
  litter.chem.phy %>%
  filter(month(sampledate) == 7)%>%
  select(lakename, year4, daynum, depth, temperature_C) %>%
  drop_na(depth,temperature_C)

#5
A6_plot1<-
  ggplot(NTL.litter, aes(x = depth, y = temperature_C)) +
  geom_point() +
  geom_smooth(method = lm, color="pink") +
  ylim(0,35)+
  ylab("Temperature in lake/Celsius")+
  xlab("Depth of lake/m")
print(A6_plot1)

```


6. Interpret the figure. What does it suggest with regards to the response of temperature to depth? Do the distribution of points suggest about anything about the linearity of this trend?

> Answer: The figure shows that as the depth of lake going deeper, the tempertature in lake will drop down. The overall distribution of points shows the negative linear relationship between the lake depth and temperature.


7. Perform a linear regression to test the relationship and display the results

```{r linear.regression}
#7
depth.tem.regression <- lm(NTL.litter$temperature_C ~ NTL.litter$depth)
summary(depth.tem.regression)
```


8. Interpret your model results in words. Include how much of the variability in temperature is explained by changes in depth, the degrees of freedom on which this finding is based, and the statistical significance of the result. Also mention how much temperature is predicted to change for every 1m change in depth. 

> Answer: In this model, 73.87% of the variability in temperature is explained by changes in depth. the degrees of freedom is 9726, the p value is much lower than 0.05 and it has statitical significant that the depth in a lake has relationship with the temperature. When there is 1m deeper in the lake depth, the temoerature degree of the lake will decrease 1.94 degree.

---

## Multiple regression
Let's tackle a similar question from a different approach. Here, we want to explore what might the best set of predictors for lake temperature in July across the monitoring period at the North Temperate Lakes LTER. 


9. Run an AIC to determine what set of explanatory variables (year4, daynum, depth) is best suited to predict temperature.

10. Run a multiple regression on the recommended set of variables. 

```{r temperature.model}
#9
temAIC <- lm(data = NTL.litter, temperature_C ~ year4 + daynum + depth)
step(temAIC)
#Here I am not sure why my model only show one option of AIC result.

temp.model <- 
  NTL.litter %>%
  select(temperature_C,year4,daynum,depth) %>%
  na.omit()
tempbyCorr <- cor(temp.model)
corrplot(tempbyCorr, method = "ellipse")
corrplot.mixed(tempbyCorr, upper = "ellipse")

#10
tem.regression <- lm(data = NTL.litter, temperature_C ~ daynum + depth)
summary(tem.regression)

```

11. What is the final set of explanatory variables that the AIC method suggests we use to predict temperature in our multiple regression? How much of the observed variance does this model explain? Is this an improvement over the model using only depth as the explanatory variable?

> Answer: Finally, the suggested set of explanatory variables should be day number and the lake depth, which could predict temperature best in our regression. There is 74.1% variability in temperature is explained by changes in depth and day number in a year. It could be an improvement over the model using only depth as the explanatory variable because the slight higher value of the adjusted R-squared. At the same time, the model got very low P value so it has statistal signiifcance that the set of depth and daynum variables has relationship with the temperature.

---
## Analysis of Variance

12. Now we want to see whether the different lakes have, on average, different temperatures in the month of July. Run an ANOVA test to complete this analysis. (No need to test assumptions of normality or similar variances.) Create two sets of models: one expressed as an ANOVA models and another expressed as a linear model (as done in our lessons).

```{r anova.model}
#12
tem.lake.anova <- aov(data = NTL.litter, temperature_C ~ lakename) 
summary(tem.lake.anova)

tem.lake.anova2 <- lm(data = NTL.litter, temperature_C ~ lakename)
summary(tem.lake.anova2)

```

13. Is there a significant difference in mean temperature among the lakes? Report your findings. 

> Answer: Yes, the p value is much lower than 0.05 so we reject the null hypothesis. The summary shows difference between a pair of lake group July mean temperature is statistically significant.



14. Create a graph that depicts temperature by depth, with a separate color for each lake. Add a geom_smooth (method = "lm", se = FALSE) for each lake. Make your points 50 % transparent. Adjust your y axis limits to go from 0 to 35 degrees. Clean up your graph to make it pretty. 

```{r,fig.height = 5, fig.width = 10, warning = FALSE}
#14.
A6_plot2<-
  ggplot(NTL.litter, aes(x = depth, y = temperature_C)) +
  geom_point(aes(color=lakename),alpha=0.5) +
  geom_smooth(method = lm,se = FALSE, fullrange=TRUE, 
              aes(color=lakename)) +
  ylim(0,35)+
  ylab("Temperature in lake/Celsius")+
  xlab("Depth of lake/m")
print(A6_plot2)

```

15. Use the Tukey's HSD test to determine which lakes have different means.

```{r tukey.test}
#15
TukeyHSD(tem.lake.anova)

```

16.From the findings above, which lakes have the same mean temperature, statistically speaking, as Peter Lake? Does any lake have a mean temperature that is statistically distinct from all the other lakes?

>Answer: Central Long and Crampton Lake;Crampton and Ward Lake;Central long lake,Tuesday lake, West long lake and East long lake have the same mean temperature with Peter lake. Hummingbird Lake have a mean temperature that is statistically distinct from all the other lakes.

 
17. If we were just looking at Peter Lake and Paul Lake. What's another test we might explore to see whether they have distinct mean temperatures? 

>Answer: Two-sample T-test could be conducted to examine the mean temperature level in two lakes.



18. Wrangle the July data to include only records for Crampton Lake and Ward Lake. Run the two-sample T-test on these data to determine whether their July temperature are same or different. What does the test say? Are the mean temperatures for the lakes equal? Does that match you answer for part 16?

```{r t.test}
Crampton.Ward <- 
  NTL.litter %>%
  filter(lakename == "Crampton Lake" | lakename == "Ward Lake")

lake.twosample <- t.test(Crampton.Ward$temperature_C ~ Crampton.Ward$lakename)
lake.twosample
  
```

>Answer:     
Ho: mean temperature in Crampton lake = mean temperature in Ward lake  
Ha: mean temperature in Crampton lake != mean temperature in Ward lake  
The results in T-test show that Pvalue is 0.2649, which is not small than 0.1 or 0.05. So we do not have enough evidence to reject the null hypothesis and we could say the mean July temperature in Crampyon lake and Ward lake are equal, which meatches with the results in question 16.
